### 2020-04-14 report/draft.tex
Report draft

\documentclass{article}

\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{soul}
\usepackage{xspace}
\usepackage{color}
\usepackage{xcolor}
\usepackage{upquote}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{wrapfig}
\usepackage{syntax}

\usepackage{tikz}
\usetikzlibrary{arrows,automata,shapes.misc,shapes.geometric,positioning}

% For comments
\newcommand{\eat}[1]{}
\newcommand{\TODO}[1]{\hl{\textbf{TODO:} #1}\xspace}
\newcommand{\todo}[1]{\hl{#1}\xspace}

%% Bibliography style
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\begin{document}

\title{CIS 673 Project Report}
\author{Gautam Mohan and Konstantinos Kallas}

\maketitle

\section{Introduction}
\label{intro}

\section{Servois}

\section{Conflict Free Replicated Data Types}

\TODO{Fill in with some background information on CRDTs}~\cite{shapiro2011conflict}

\section{Methodology}

\section{Experience Report}

%% Bibliography
\bibliography{./draft}
\bibliographystyle{plain}

\end{document}